Non-word repetition in Yélî Dnye

### 2020-09-14 manuscript.Rmd
addressed middys comments code review

abstract: >
   In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.

@@

- look through this manuscript for bits of text to be written by **mc**
- do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check phonological implementation of stimuli in stimuli.xlsx 
- for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
- double check also stim_seg_freq -- why are some items still with NA freq?
- consent, "correct", "accuracy", "performance"
- draft discussion

@@

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood [e.g., @hazan2000development]. One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. Accuracy is thought to reflect long-term phonological knowledge (which allows the child to perceive the item accurately even though it is not a real word they have encountered before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item accurately even though it had never been pronounced before).

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level accuracy [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being "bilik", and imagine a response "bilig". This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20% (20% of phonemes were substituted or deleted).  
<!--M2A: A concrete scoring example would be a good idea here-->
<!-- a2m added! actually, I haven't changed anything but I've been hesitant about using levenshtein's distance because it's the complement of phoneme based scoring -- so shall we drop it?  -->

@@

#quick fix, but should be fixed upstream @ac
# M2A: Seems like you already did this, and the line below
# is setting the is.na(trnsc$Sex) cases to "G" -- I assume
# that's not a desirable outcome so am commenting it out
# trnsc$Sex[trnsc$Sex=="F"]<-"G"
trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)
```

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. We excluded `r excl` children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt!="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt!="first"]))`)
were on average less accurate than first ones (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt=="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt=="first"]))`), t(`r round(myt$parameter,2)`) = `r round(myt$statistic,2)`, p = `r round(myt$p.value,2)`. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,pc$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

@@

In this analysis, we were interested in variation in performance as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID & target ID as random variables.

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings [@casillas2020png], and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(`r freq_cor$parameter`)=`r round(freq_cor$estimate,3)`, =`r round(freq_cor$p.value,3)`). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = `r both_coeff["avg_fr_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_s","Std. Error"]`, p = `r both_coeff["avg_fr_s","Pr(>|z|)"]`) and age (ß = `r both_coeff["age.rounded_s","Estimate"]`, SE ß = `r both_coeff["age.rounded_s","Std. Error"]`, p = `r both_coeff["age.rounded_s","Pr(>|z|)"]`) were significant predictors of accuracy, but corpus frequency (ß = `r both_coeff["avg_fr_cor_ty_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_cor_ty_s","Std. Error"]`, p = `r both_coeff["avg_fr_cor_ty_s","Pr(>|z|)"]`) was not.

@@

### Accuracy as a function of item length

```{r bylength}
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->by_length
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->nld
by_length=cbind(by_length,pc$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

Next, we inspected whether accuracy varied as a function of word length. Results are shown on table XX. This table shows that monosyllables accuracy was much lower than other lengths. This is likely due to the fact that the majority of monosyllables included were chosen because they had sounds that are rare in the world's languages, which may indicate that they are hard to produce or to perceive.

Setting monosyllables aside, we observe the typical pattern of decreased accuracy for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors.

```{r tablength}
birthorder.tab=NULL
for(myvar in c("Word","Phoneme","NLD")){
  birthorder.tab=cbind(birthorder.tab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(birthorder.tab)<-levels(by_length$Length)
colnames(birthorder.tab)<-c("Word","Phoneme","NLD")
apa_table(birthorder.tab,caption="NWR scores measured in whole word accuracy, phoneme-based accuracy, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```

@@

- What is the overall repetition accuracy (whole word, phoneme based, distance)?

@@

1. Children are more accurate for mono-syllables than longer items
1. The length distribution in Yélî words is more balanced than that in English, and thus the performance decline for poly- versus mono-syllables may be less pronounced than that for English. **Check for work on European languages that may have looked into this**

2. Similarly, we do not know of NWR research that manipulates the difficulty of the sounds that are included in the items, but word naming and other research suggests that children are more accurate when producing easy and/or typologically common sounds than difficult and/or typologically rare sounds [CITE]. Therefore, we expect higher accuracy for items with common sounds than in those with rare sounds.
2. The Yélî sound inventory is very large and compressed, with many similar sounds that are acoustic and articulatory neighbors. Therefore, this may constitute a pressure for children to have finer auditory skills (and perhaps more precise articulations) than children speaking languages with a simpler inventory. As a result, differences between easier and harder items may be smaller in this work than in other research. **no work looking at consonants & vowels? no work looking at nasal vowels in particular?**

@@

3. Children's accuracy increases with child age.
4. Non-monolingual Yélî children are less accurate than monolingual ones when tested on the society-dominant language (we did not test any non-dominant language) 
<!-- **(MC: or just because they're acquiring two phonologies and so show slightly delayed linguistic development, in parallel with vocab development stuff; we don't have data on their dominant language, but i guess we can assume it's YD for most kids---i'm sure there are exceptions!)** ac2mc I added "society-" because that's what I meant (we didn't get proportions of time in which they hear YD versus other languages); I didn't add anything about why this would come out like this or not here, but I'll add this -- although I think the evidence on that is mixed (ie no strong evidence that learning 2 phonologies sets you back)-->
5. As revious NWR evidence on this is mixed, but general findings on language development suggest that children whose mothers are more educated are more accurate than children whose mothers are less educated.
6. To our knowledge, there is no previous NWR work on this, but other research suggests that first-born children should outperform later-born children

4. Anecdotally Yélî children grow up in close-knitted communities and thus may receive significant portions of their language input from people not in their nuclear family 
<!-- **(MC: definitely almost all kin; 5k of them)** ac2mc I'm not sure this is how bio anthropologists think about kin; because it's about how you attribute resources to boost your genes' chances of being passed on relative to those of others in the same space. So my guess is that there is still a difference between nuclear family and not. Rewrote kin -> nuclear family -->
(or at least from people other than their mothers, who tend to be the non-native speakers). If so, the difference between monolinguals and not monolinguals may be smaller than that found in other work
<!-- **(MC: maybe it'd be more useful to try and estimate the number of non-Y moms who live in each hamlet where testing is taking place?)** ac2mc not sure it's worth the effort, in any case other female adults are almost certain to have very little impact on the target child... Perhaps I'm not saying something very clearly? -->
. That said, one recent study on the same population shows that most child-directed input in the first 2 years does come from the mother 
<!-- **(MC: i can get you actual estimates for this for YD; let's go with that instead!)** -->
<!-- ac2mc I think I meant your own paper! added "on the same population "-->
, so in so far as this input has a crucial formational role, then there may still be a performance gap between these two groups.
5. In the Rossel community, formal education plays an extremely minor role in ensuring individual's success, is not a good index of relative socio-economic status, and furthermore there is only a narrow range of variation in maternal educational attainment. This may lead to no or only very small advantages for children whose mothers are more educated, provided that the causal chain between maternal education and child language is via SES more broadly. However, if education directly boosts maternal verbal skills and the incidence of verbal behavior (as suggested by CITE), then we should still see a difference along this factor.
6. One main causal path between birth order and language development is via parental input (CITE). Given our arguments above for how mothers may not be as important among Rossel people than in other places, then the performance gap between first borns and later borns may be smaller.

>>>>

abstract: >
   In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. NWR scores are thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated without changes less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.

@@

- look through this manuscript for bits of text to be written by **mc**
- do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check phonological implementation of stimuli in stimuli.xlsx 
- for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
- double check also stim_seg_freq -- why are some items still with NA freq?
- draft discussion

## TODO Alex

- add effect sizes to all analyses
- reply to questions

@@

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood [e.g., @hazan2000development]. One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. NWR scores are thought to reflect long-term phonological knowledge (which allows the child to perceive the item precisely even though it is not a real word they have encountered before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely even though it had never been pronounced before).

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level scoring [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being "bilik", and imagine a response "bilig". This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20% (20% of phonemes were substituted or deleted).  
<!--M2A: A concrete scoring example would be a good idea here-->
<!-- a2m added! actually, I haven't changed anything but I've been hesitant about using levenshtein's distance because it's the complement of phoneme based scoring -- so shall we drop it?  -->

@@

#quick fix, but should be fixed upstream @ac
# M2A: Seems like you already did this, and the line below
# is setting the is.na(trnsc$Sex) cases to "G" -- I assume
# that's not a desirable outcome so am commenting it out
# a2m fixed, but I think it is better to fix all of these error upstream, in the demo.xlsx file
trnsc$Sex[trnsc$Sex=="F" & !is.na(trnsc$Sex)]<-"G"
trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)
```

@@

We first checked whether NWR scores varied between first and subsequent presentations of an item by averaging word-level scores at the participant level separately for first attempts and subsequent repetitions. We excluded `r excl` children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level scores became more heterogeneous in subsequent repetitions. Surprisingly, NWR scores for subsequent repetitions (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt!="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt!="first"]))`)
were on average lower than first ones (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt=="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt=="first"]))`), t(`r round(myt$parameter,2)`) = `r round(myt$statistic,2)`, p = `r round(myt$p.value,2)`. Given the uncertainty in whether previous work used only the first or all repetitions, and since scores declined and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,phon_score$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

@@

In this analysis, we were interested in variation in NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID & target ID as random variables.

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings [@casillas2020png], and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(`r freq_cor$parameter`)=`r round(freq_cor$estimate,3)`, =`r round(freq_cor$p.value,3)`). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = `r both_coeff["avg_fr_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_s","Std. Error"]`, p = `r both_coeff["avg_fr_s","Pr(>|z|)"]`) and age (ß = `r both_coeff["age.rounded_s","Estimate"]`, SE ß = `r both_coeff["age.rounded_s","Std. Error"]`, p = `r both_coeff["age.rounded_s","Pr(>|z|)"]`) were significant predictors of word-level NWR scores, but corpus frequency (ß = `r both_coeff["avg_fr_cor_ty_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_cor_ty_s","Std. Error"]`, p = `r both_coeff["avg_fr_cor_ty_s","Pr(>|z|)"]`) was not.

@@

### NWR scores as a function of item length

```{r bylength}
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->by_length
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->nld
by_length=cbind(by_length,phon_score$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

Next, we inspected whether NWR scores varied as a function of word length. Results are shown on table XX. This table shows that scores for monosyllables was much lower than for other lengths. This is likely due to the fact that the majority of monosyllables included were chosen because they had sounds that are rare in the world's languages, which may indicate that they are hard to produce or to perceive.

Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors.

```{r tablength}
birthorder.tab=NULL
for(myvar in c("Word","Phoneme","NLD")){
  birthorder.tab=cbind(birthorder.tab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(birthorder.tab)<-levels(by_length$Length)
colnames(birthorder.tab)<-c("Word","Phoneme","NLD")
apa_table(birthorder.tab,caption="NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```

@@

- What is the overall repetition scores (whole word, phoneme based, distance)?

@@

1. Prediction based on previous work: Children have higher scores for shorter items
1. Prediction for Yélî made before seeing the data: The length distribution in Yélî words is more balanced than that in English, and thus the score decline for poly- versus mono-syllables may be less pronounced than that for English. 
1. it turns out we were right! the scores for 2-4 syll words decline only slightly with length

2. Prediction based on previous work: Similarly, we do not know of NWR research that manipulates the difficulty of the sounds that are included in the items, but word naming and other research suggests that children achieve higher scores when producing easy and/or typologically common sounds than difficult and/or typologically rare sounds [CITE]. Therefore, we expect higher scores for items with common sounds than in those with rare sounds.
2. Prediction for Yélî made before seeing the data: The Yélî sound inventory is very large and compressed, with many similar sounds that are acoustic and articulatory neighbors. Therefore, this may constitute a pressure for children to have finer auditory skills (and perhaps more precise articulations) than children speaking languages with a simpler inventory. As a result, differences between easier and harder items may be smaller in this work than in other research. 
2. it turns out we do see

@@

3. Children's scores increases with child age.

4. General prediction: Non-monolingual Yélî children have lower scores than monolingual ones when tested on the society-dominant language (we did not test any non-dominant language) 
<!-- **(MC: or just because they're acquiring two phonologies and so show slightly delayed linguistic development, in parallel with vocab development stuff; we don't have data on their dominant language, but i guess we can assume it's YD for most kids---i'm sure there are exceptions!)** ac2mc I added "society-" because that's what I meant (we didn't get proportions of time in which they hear YD versus other languages); I didn't add anything about why this would come out like this or not here, but I'll add this -- although I think the evidence on that is mixed (ie no strong evidence that learning 2 phonologies sets you back)-->
4. Local prediction: Anecdotally Yélî children grow up in close-knitted communities and thus may receive significant portions of their language input from people not in their nuclear family 
<!-- **(MC: definitely almost all kin; 5k of them)** ac2mc I'm not sure this is how bio anthropologists think about kin; because it's about how you attribute resources to boost your genes' chances of being passed on relative to those of others in the same space. So my guess is that there is still a difference between nuclear family and not. Rewrote kin -> nuclear family -->
(or at least from people other than their mothers, who tend to be the non-native speakers). If so, the difference between monolinguals and not monolinguals may be smaller than that found in other work
<!-- **(MC: maybe it'd be more useful to try and estimate the number of non-Y moms who live in each hamlet where testing is taking place?)** ac2mc not sure it's worth the effort, in any case other female adults are almost certain to have very little impact on the target child... Perhaps I'm not saying something very clearly? -->
. That said, one recent study on the same population shows that most child-directed input in the first 2 years does come from the mother 
<!-- **(MC: i can get you actual estimates for this for YD; let's go with that instead!)** -->
<!-- ac2mc I think I meant your own paper! added "on the same population "-->
, so in so far as this input has a crucial formational role, then there may still be a score difference between these two groups.
4. We don't see a difference

5.  General prediction: previous NWR evidence on this is mixed, but general findings on language development suggest that children whose mothers are more educated score higher in standardized language tests (eg ppvt) than children whose mothers are less educated. 
5. Prediction made before seeing data: In the Rossel community, formal education plays an extremely minor role in ensuring individual's success, is not a good index of relative socio-economic status, and furthermore there is only a narrow range of variation in maternal educational attainment. This may lead to no or only very small advantages for children whose mothers are more educated, provided that the causal chain between maternal education and child language is via SES more broadly. However, if education directly boosts maternal verbal skills and the incidence of verbal behavior (as suggested by CITE), then we should still see a difference along this factor.
5. We did not find this

6. General prediction: To our knowledge, there is no previous NWR work on this, but other research suggests that first-born children score higher in standardized language tests later-born children. 
6. Prediction before seeing data: One main causal path between birth order and language development is via parental input (CITE). Given our arguments above for how mothers may not be as important among Rossel people than in other places, then the difference in scores between first borns and later borns may be small
6. we did not find a sig effect, but this is a small effect (eg d=.2 in Havron et al)
